---
title: "GGplot Chart Theme Sampler"
format: html
editor: visual
execute: 
  echo: true
  warning: false
---

```{r setup, include = FALSE}
library(tidyverse)
library(scales)
library(lubridate)
library(RColorBrewer)

```

## Charts with different ggplot themes 

Designed to help with choosing the best option for the situation.

Uses themes that come with the ggplot package.

### First, get some data

```{r get_data}
# data set that comes with R and has time dimension and at least one other dimension and at least one metric
chart_data <- tsibbledata::global_economy
# add metric for GDP per capita
chart_data <- chart_data %>% mutate(
  GDP_per_cap = GDP / Population
)
# isolate country for use case
chart_data_single <- chart_data %>% filter(Code=='CAN')
# filter for select countries to make manageable for use cases
chart_data_select <- chart_data %>% filter(Code %in% c("CAN","USA","BRA","JPN","IND","NGA")) 
```

### Function to take chart info and process themes

```{r chart_themes}
# function to take chart info as input, apply themes, return variations with themes
chartThemeSampler <- function(chart_example) {
  c_dflt <- chart_example + labs(title = "Default (aka Gray/Grey)")
  c_m <- chart_example + labs(title = "Minimal") + theme_minimal()
  c_l <- chart_example + labs(title = "Light") + theme_light()
  c_drk <- chart_example + labs(title = "Dark") + theme_dark()
  c_c <- chart_example + labs(title = "Classic") + theme_classic()
  c_bw <- chart_example + labs(title = "BW") + theme_bw()
  return(list(c_dflt, c_m, c_l, c_drk, c_c, c_bw))
}
```

## Histograms

Using histogram as example to show general look/feel of different themes.

Separated here with headings for clarity. Subsequent sections theme is noted in chart title.

```{r histogram}
c1 <- chart_data_select %>% ggplot(aes(x=GDP_per_cap))+geom_histogram()
```

### Default Chart (aka Gray/Grey)

```{r}
c1
```

### Minimal

```{r}
c1 + theme_minimal()
```

### Light

```{r}
c1 + theme_light()
```

### Dark

```{r}
c1 + theme_dark()
```

### Classic

```{r}
c1 + theme_classic()
```

### BW

```{r}
c1 + theme_bw()
```

## Line Charts

```{r}
# set chart for target style
chart_example <- chart_data_single %>% ggplot(aes(x=Year, y=Growth))+geom_line()
# print charts in different themes
chartThemeSampler(chart_example)
```

## Column (Bar) Charts

```{r}
# set chart for target style
chart_example <- chart_data_single %>% ggplot(aes(x=Year, y=Growth))+geom_col()
# print charts in different themes
chartThemeSampler(chart_example)
```

### Column Charts with Faceting

```{r}
chart_example <- chart_data_select %>% filter(Year >= 2000) %>%
  ggplot(aes(x=Year, y=Growth))+geom_col()+
  facet_grid(Code~.)
chartThemeSampler(chart_example)
```

## Boxplots

```{r}
chart_example <- chart_data_select %>% 
  ggplot(aes(x=Code, y=Growth))+geom_boxplot()
chartThemeSampler(chart_example)
```
